|

### 2020-02-02 analysis/paper/paper.Rmd
Discuss how to identify and deal with violations of the single encounter assumption.

\begin{abstract}
Capture-recapture studies are widely used in ecology to estimate population sizes and demographic rates. 
In some capture-recapture studies, individuals may be visually encountered but not identified. 
For example, if individual identification is only possible upon capture and individuals escape capture, visual encounters can result in failed captures where individual identities are unknown.
In such cases, the data consist of capture histories with known individual identities, and counts of failed captures for individuals with unknown identities.
These failed captures are ignored in traditional capture-recapture analyses that require known individual identities.
Here we show that if animals can be encountered at most once per sampling occassion, failed captures provide lower bounds on population size that can increase the precision of abundance estimates. 
Analytical results and simulations indicate that visual encounter data improve abundance estimates when capture probabilities are low, and when there are few repeat surveys. 
We present a hierarchical Bayesian approach for integrating failed captures and auxiliary encounter data in statistical capture-recapture models. 
This approach can be integrated with existing capture-recapture models, and may prove particularly useful for hard to capture species in data-limited settings.
\end{abstract}

@@

Capture-recapture studies are widely used for estimating abundance and demographic rates, using information about the identities of captured individuals [@jolly1965explicit]. 
This paper examines the case where individual identification requires capture, and identities of animals that are visually encountered but not captured are unknown.
In such cases, "capture" is synonymous with "identification". 
This often applies in capture-recapture studies of amphibians, where individual identification is only possible with the animal in hand. 
We also assume that if an individual escapes capture, it is not encountered again in the sampling occassion because it is hiding or otherwise inaccessible [@gibbons2004pit; @bailey2010capture; @joseph2018disease]. 
Under these conditions, encounters leading to failed captures provide information about abundance, but this information is not readily used in traditional capture-recapture models.

@@

Here, we show how visual encounter data can improve abundance estimates in capture-recapture studies for study designs where 1) individual identification requires capture (so that a subset of encountered individuals are captured), and 2) individuals can be encountered at most once per sampling occassion. 
We develop a modified capture-recapture observation model, and investigate conditions under which encounter data improve abundance estimates.
The methods presented here accommodate both failed captures and counts of animals collected separately from capture-recapture surveys, and can be integrated with existing capture-recapture models.

@@

Here we assumed that individuals could be encountered at most once per sampling occassion. 
In other words, there is no double counting: each failed capture corresponds to one unique individual. 
The observation model developed here is not robust to violations of this assumption, because the total number of encounters sets a lower bound on the true population size. 
As a consequence, if individuals escape capture multiple times in the sampling sampling occassion, it is possible that the posterior for abundance might be misleadingly precise (i.e., the lower bound on population size would be too high).
Therefore, we do not recommend this approach if individuals might be encountered multiple times on the same sampling occassion. 
This assumption is likely to hold for example in capture-recapture studies of amphibians in high elevation lakes, where individual animals that escape capture hide afterwards, e.g., underwater where they cannot be seen or captured [@joseph2018disease].

These results are consistent with related findings for mark-resight studies where marked individuals are subject to incomplete identification. 
In such studies, accounting for failed identifications of marked individuals -- analogous to failed captures -- is most advantageous when identification probabilities are low -- analogous to capture probabilities being low [@mcclintock2014mark]. 
However, in the encounter-capture-recapture scenario considered here, we assume that whether an animal is marked or not is unknown until it is captured (as would be the case for subdermal PIT tags in an amphibian).

>>>>

\begin{abstract}
Capture-recapture studies are widely used in ecology to estimate population sizes and demographic rates. 
In some capture-recapture studies, individuals may be visually encountered but not identified. 
For example, if individual identification is only possible upon capture and individuals escape capture, visual encounters can result in failed captures where individual identities are unknown.
In such cases, the data consist of capture histories with known individual identities, and counts of failed captures for individuals with unknown identities.
These failed captures are ignored in traditional capture-recapture analyses that require known individual identities.
Here we show that if animals can be encountered at most once per sampling occasion, failed captures provide lower bounds on population size that can increase the precision of abundance estimates. 
Analytical results and simulations indicate that visual encounter data improve abundance estimates when capture probabilities are low, and when there are few repeat surveys. 
We present a hierarchical Bayesian approach for integrating failed captures and auxiliary encounter data in statistical capture-recapture models. 
This approach can be integrated with existing capture-recapture models, and may prove particularly useful for hard to capture species in data-limited settings.
\end{abstract}

@@

Capture-recapture studies are widely used for estimating abundance and demographic rates, using information about the identities of captured individuals [@jolly1965explicit]. 
This paper examines the case where individual identification requires capture, and identities of animals that are visually encountered but not captured are unknown.
In such cases, "capture" is synonymous with "identification". 
This often applies in capture-recapture studies of amphibians, where individual identification is only possible with the animal in hand. 
We also assume that if an individual escapes capture, it is not encountered again in the sampling occasion because it is hiding or otherwise inaccessible [@bailey2010capture; @joseph2018disease]. 
Under these conditions, encounters leading to failed captures provide information about abundance, but this information is not readily used in traditional capture-recapture models.

@@

Here, we show how visual encounter data can improve abundance estimates in capture-recapture studies for study designs where 1) individual identification requires capture (so that a subset of encountered individuals are captured), and 2) individuals can be encountered at most once per sampling occasion. 
We develop a modified capture-recapture observation model, and investigate conditions under which encounter data improve abundance estimates.
The methods presented here accommodate both failed captures and counts of animals collected separately from capture-recapture surveys, and can be integrated with existing capture-recapture models.

@@

Here we assumed that individuals could be encountered at most once per sampling occasion. 
In other words, there is no double counting: each failed capture corresponds to one unique individual. 
The observation model developed here is not robust to violations of this assumption, because the total number of encounters sets a lower bound on the true population size. 
As a consequence, if individuals escape capture multiple times in the sampling sampling occasion, it is possible that the posterior for abundance might be misleadingly precise (i.e., the lower bound on population size would be too high).
Therefore, we do not recommend this approach if individuals might be encountered multiple times on the same sampling occasion. 
This assumption is likely to hold for example in capture-recapture studies of amphibians in high elevation lakes, where individual animals that escape capture hide afterwards, e.g., underwater where they cannot be seen or captured [@joseph2018disease]. 
If individuals are captured multiple times in the same sampling occasion, then it will be clear that the assumption of at most one encounter has been violated. 
In such cases, alternative encounter models may be necessary, e.g., a Poisson model that allows repeat encounters on a sampling occasion [@royle2009bayesian].

These results are consistent with related findings for mark-resight studies where marked individuals are subject to incomplete identification. 
In such studies, accounting for failed identifications of marked individuals -- analogous to failed captures -- is most advantageous when identification probabilities are low -- analogous to capture probabilities being low [@mcclintock2014mark]. 
However, in the encounter-capture-recapture scenario considered here, we assume that whether an animal is marked or not is unknown until it is captured, as would be the case for subdermal passive integrated transponder tags in an amphibian [@gibbons2004pit].
